\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-14 k-planar.tex
Touch ups

@@

For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$.
% \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. }
The graph $G_0$ is called the \defin{frame} of $G_0$. This definition (for simple graphs) was introduced by \citet{BDGGMR}. Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph. Thus \cref{ShortcutProduct} is applicable. However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}.

@@

This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs. For simple 1-planar graphs it is essential that we allow frames with parallel edges.

\cref{d_framed_product_stucture} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}. The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.  Finally, a \emph{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\dist_G(v,V_0)=\dist_T(v,V_0)$ for each $v\in V(G)$.

\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
		\item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
		\begin{compactenum}
			\item if $k\in\{2,3\}$ then $F[P_i]$ is a path; and
			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
		\end{compactenum}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
	\begin{compactenum}[(i)]
		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
	\end{compactenum}
\end{lem}

@@

	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs when $V(N')=V(F)$, for which we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
	\begin{compactenum}[(i)]
		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
		\item Assumption 4(a) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
	\end{compactenum}

	In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.  There is one case to consider for each possible value of $k$:
	\begin{compactenum}
		\item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.

@@

	\begin{figure}
		\begin{center}
			\begin{tabular}{ccc}
				\includegraphics{figs/zoomba-1} &
				\includegraphics{figs/zoomba-2} &
				\includegraphics{figs/zoomba-3} \\
				$k=1$ & $k=2$ & $k=3$
			\end{tabular}
		\end{center}
		\caption{Three cases when choosing the face $\tau$ of $N'$.}
		\label{boring_figure}
	\end{figure}

	Refer to \cref{boring_figure}.  Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$. By our choice of $\tau$, the graph $M$ is $2$-connected.  Since $N'\subseteq G'$ contains only edges of $G$ that are not crossed, each component of $G-V(M)$ is contained in a single face of $M$.
	Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$. By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.

	In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap Y\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.  When this assumption does not hold, exactly the same reasoning can be used, while omitting one or more of $P_{i,1}$, $P_{i,2}$ or $P_{i,3}$.

	Since $M$ is $2$-connected and $\overline{Y}$ is connected, $F_i[P_{i,j}]$ is a path, for each $j\in\{1,2,3\}$.  Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.  Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.  Therefore $N_i'$ has fewer inner faces than $N'$.  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.  The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of the subgraph $N_i$ of $N$ contained in the closure of $F_i$.
	% The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
	We can now define our desired partition
	\[
	\mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
	\]
	It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.  All that remains is to verify that $\mathcal{P}$ satisfies (i)--(iii).

@@

	All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii). Let $x_0$ be the vertex of $H$ obtained from contracting $S$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.  The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$. The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).  Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively. Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ contains $\{a_i,b_i,c_i\}$.

	Recall that $P_{i,1}:=V(N_i)\cap Y$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.  For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.  These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$. Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.  We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $mathcal{T}:=(B_y:y\in V(T_0))$.

  It is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $H':=N'/\mathcal{P}$ in which each bag has size at most $4$.    To see that $\mathcal{T}$ is also a tree decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N')$ with $v\in P_\alpha$ and $w\in P_\beta$.  We must verify that some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$.  If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.  If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $G$ crosses an edge of $M$, $v$ or $w$ are both contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.  Finally, if $v\in V(M)$ and $w\not\in V(M)$ then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$. By induction, some bag of $\mathcal{T}_i$ contains $\beta$ and $\alpha'$, where $\alpha'$ is the vertex of $H_i$ obtained by contracting $P_{i,j}$.  However, before $T_i$ is attached to $T_0$ each occurence of $\alpha'$ in $\mathcal{T}_i$ is replaced by $\alpha$, so some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$. is renamed to $\alpha$.  Therefore $\mathcal{T}$ is a tree decomposition of $H$.

  Next we verify that, when $k=3$ excactly one bag of $T$ contains $\{x_1,\ldots,x_3\}$.  The bag $B_{y_0}$ contains $x_1,\ldots,x_k$. When $k=3$, each neighbour of $z_i$ of $y_0$ has a bag that contains at most two of $\{x_1,x_2,x_3\}$.  Therefore, when $k=3$, exactly one bag of $\mathcal{T}$ contains $\{x_1,x_2,x_3\}$.

	All that remains is to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$. By the inductive hypothesis, the only triples we need to be concerned about are the $3$-element subsets of  $\{x_0,x_1,\ldots,x_k\}$. The case $k=1$ is trivial since no triples can be formed by $\{x_0,x_1\}$.  When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.  In this case, the choice of $\tau$ ensures that there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_1$, and $P_2$ on its boundary.  (Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.  Therefore there are at most two bags of $\mathcal{T}$ that contains $\{x_0,x_1,x_2\}$.  This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.

	We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.  Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.  Create a graph $M^+$ from $M$ by placing a vertex $v_{i_b}$ inside $F_{i_b}$ and adjacent to each vertex in $V(F_{i_b})$ for each $b\in\{1,2\}$.  Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.  However, contracting each of $S_{x_0},\ldots,S_{x_3}$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.  (In this contracted graph, $x_0,x_a,x_b$ are on one side of the bipartition and $v_{i_a}$, $v_{i_b}$, and the node in $\{x_1,x_2,x_3\}\setminus\{x_a,x_b\}$ is on the other side.)

	We have already established that the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ with $1\le a<b\le 3$. By the preceding claim, there is at most one face $F_i$ with vertices of $S$, $P_a$, and $P_b$ on its boundary.  By the inductive hypothesis, the only other bag in $\mathcal{T}$ that contains $\{x_0,x_a,x_b\}$ is the bag $B_{z_i}$.  This completes the proof.
\end{proof}

@@

	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdot \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$. That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.  Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$. The result follows from \cref{PartitionProduct}.
\end{proof}

>>>>

\definecolor{brew2}{rgb}{1.0, 1.0, 0.701960784314}
\definecolor{brew8}{rgb}{0.988235294118, 0.803921568627, 0.898039215686}

@@

For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that the interior of each edge in $E(G)\setminus E(G_0)$ is contained in the interior of some face of $G_0$.
% \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. }
The graph $G_0$ is called the \defin{frame} of $G_0$. This definition (for simple graphs) was introduced by \citet{BDGGMR}. Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph. Thus \cref{ShortcutProduct} is applicable. However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible.\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}

@@

This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs. \todo{Move this comment to the section on $1$-planar graphs?} For simple 1-planar graphs it is essential that we allow frames with parallel edges.

\cref{d_framed_product_stucture} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}. The following definition is a convenient way to establish the planarity of $H$ in \cref{d_framed_product_stucture}:  A tree-decomposition $\mathcal{T}$ of a graph $G$ is \defin{$t$-simple} if each of its bags has size at most $t+1$ and, for each $t$-element subset $\{v_1,\ldots,v_t\}\subseteq V(G)$, at most two bags of $\mathcal{T}$ contain $\{v_1,\ldots,v_t\}$.  Finally, a \defin{BFS spanning forest} $T$ of $G$ rooted at a set $V_0\subseteq V(G)$ is a spanning forest of $G$ with the property that $\min\{\dist_G(v,w):w\in V_0\}=\min\{\dist_T(v,w):w\in V_0\}$ for each $v\in V(G)$.

\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
		\item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
		\item Let $F$ be a cycle in $G'$ that $V(F)$ is partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
		\begin{compactenum}
			\item $F[P_i]$ is connected; and
			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
		\end{compactenum}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
	\begin{compactenum}[(i)]
		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
	\end{compactenum}
\end{lem}

@@

	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs when $V(N')=V(F)$, for which we can simply take $\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
	\begin{compactenum}[(i)]
		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
		\item Assumption 4(b) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
	\end{compactenum}

	In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that is defined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.  There is one case to consider for each possible value of $k$ (see \cref{boring_figure}):
	\begin{compactenum}
		\item If $k=1$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge $v_1v_2$ of $F$ on its boundary.

@@

	\begin{figure}
		\begin{center}
			\begin{tabular}{c@{}c@{}c}
				\includegraphics{figs/zoomba-1} &
				\includegraphics{figs/zoomba-2} &
				\includegraphics{figs/zoomba-3} \\
        \multicolumn{3}{c}{{\color{brew8}\raisebox{-3pt}{\rule{12pt}{12pt}}} $S$\qquad {\color{brew2}\raisebox{-3pt}{\rule{12pt}{12pt}}} $X$} \\
				$k=1$ & $k=2$ & $k=3$
			\end{tabular}
		\end{center}
		\caption{Three cases when choosing the face $\tau$ of $N'$.}
		\label{boring_figure}
	\end{figure}

	Refer to \cref{boring_figure}.  Let $M$ denote the subgraph of $N'$ containing the edges and vertices of $\overline{S}$ and the edges and vertices of $F$. By our choice of $\tau$, the graph $M$ is $2$-connected.  
  % Since $N'\subseteq G'$ contains only edges of $N\subseteq G$ that are not crossed, each component of $N-V(M)$ is contained in a single face of $M$.
	Let $F_1,\ldots,F_m,\tau$ be the inner faces of $M$. By our choice of $v_1v_2v_3$, the vertices of each $F_i$ can be partitioned into at most three sets $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$ where $P_{i,1}\subset Y$, $P_{i,2}\subseteq P_b$, and $P_{i,3}\subseteq P_c$ for some $b,c\in\{1,\ldots,k\}$.

	In the following, in order to avoid introducing even more notation or constantly adding the modifier ``when it exists'', we will assume that $V(F_i)\cap S\neq\emptyset$ and that there are exactly two distinct $b,c\in\{1,\ldots,k\}$ such that $V(F_i)\cap P_b\neq\emptyset$ and $V(F_i)\cap P_c\neq\emptyset$.  When this assumption does not hold, exactly the same reasoning can be used, while omitting one or more of $P_{i,1}$, $P_{i,2}$ or $P_{i,3}$.

	Since $M$ is $2$-connected and $\overline{S}$ is connected, $F_i[P_{i,j}]$ is connected, for each $j\in\{1,2,3\}$.  Let $N_i$ and $N_i'$ be the subgraphs of $N$ and $N'$ contained in the closure of $F_i$.  Every inner face of $N_i'$ is an inner face of $N'$ and $\tau$ is not an inner face of $N_i'$.  Therefore $N_i'$ has fewer inner faces than $N'$.  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_{i,1},P_{i,2},P_{i,3}$ of $V(F_i)$.  The result of the induction is a $H_i$-partition $\mathcal{P}_i$ of $N_i$.
	% The partition $\mathcal{P}_i$ contains each of $P_{i,1}$, $P_{i,2}$ and $P_{i,3}$.
	We can now define our partition
	\[
	\mathcal{P}:=\{Y, P_1,\ldots,P_k\} \cup \bigcup_{i=1}^m \{P\in\mathcal{P}_i: P\cap V(M)=\emptyset\} \enspace .
	\]
	It is straightforward to verify that $\mathcal{P}$ is indeed a partition of $V(N)$ and, by definition this is an $H$-partition of $N$ for the graph $H:=N/\mathcal{P}$.  All that remains is to verify that $\mathcal{P}$ satisfies (i)--(iii).

@@

	All that remains is to construct a tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$ of $H$ that satisfies (iii). Let $x_0$ be the vertex of $H$ obtained from contracting $S$ and, for each $j\in\{1,\ldots,k\}$, let $x_j$ denote the vertex of $H$ obtained from contracting $P_j$.  The tree $T_0$ has a node $y_0$ with bag $B_{y_0}:=\{x_0,\ldots,x_k\}$. The inductive hypothesis ensures that, for each $i\in\{1,\ldots,m\}$,  $H_i:=N_i/\mathcal{P}_i$ has a tree decomposition $\mathcal{T}_i:=(B_z:z\in V(T_i))$ that satisfies (iii).  Let $a_i,b_i,c_i\in V(H_i)$ be the vertices obtained by contracting $P_{i,1}$, $P_{i,2}$, and $P_{i,3}$, respectively. Since $a_i$, $b_i$, $c_i$ form a clique in $H$, there exists a bag $B_{z_i}$ in $\mathcal{T}_i$ that contains $\{a_i,b_i,c_i\}$.

	Recall that $P_{i,1}:=V(N_i)\cap S$, $P_{i,2}:=V(N_i)\cap P_b$ and $P_{i,3}:=V(N_i)\cap P_c$ for some $b,c\in\{1,\ldots,k\}$.  For each bag $B_z$, $z\in V(T_i)$, replace each occurrence of $a_i$ by $x_0$, replace each occurrence of $b_i$ by $x_b$ and replace each occurrence of $c_i$ by $x_c$.  These replacements do not increase the size of any bag and they result in a tree decomposition of the graph obtained from $H_i$ by renaming the vertices $a_i$, $b_i$, and $c_i$ with the names $x_0$, $x_a$, and $x_b$. Now add an edge that joins $y_0$ to $z_i$ so that $T_i$ becomes a subtree of $T_0$ that is adjacent to $y_0$.  We perform this operation for each $i\in\{1,\ldots,m\}$ to obtain our final tree decomposition $\mathcal{T}:=(B_y:y\in V(T_0))$.

  It is now straightforward to verify that $\mathcal{T}$ is a tree decomposition of $H':=N'/\mathcal{P}$ in which each bag has size at most $4$.    To see that $\mathcal{T}$ is also a tree decomposition of $H:=N/\mathcal{P}$, consider any edge $vw\in E(N)\setminus E(N')$ with $v\in P_\alpha$ and $w\in P_\beta$.  We must verify that some bag of $\mathcal{T}$ contains $\alpha$ and $\beta$.  If both $v$ and $w$ are in $V(M)$, then $\alpha=x_a$ and $\beta=x_b$ for some $a,b\in\{0,\ldots,3\}$, in which case $B_{y_0}=\{x_0,\ldots,x_k\}$ contains $\alpha$ and $\beta$.  If neither $v$ nor $w$ are in $V(M)$ then, since no edge of $N\subseteq G$ crosses an edge of $M\subseteq G'$, both $v$ and $w$ are contained in the interior of the same face $F_i$ of $M$ so, by the inductive hypothesis there exists a bag of $\mathcal{T}_i$ that contains $\alpha$ and $\beta$ and this becomes a bag of $\mathcal{T}$ that contains $\alpha$ and $\beta$.  Finally, if $v\in V(M)$ and $w\not\in V(M)$ then $v\in P_{i,j}$ for some $j\in\{1,2,3\}$ and $w$ is in the interior of some face $F_i$ of $M$. By induction, some bag of $\mathcal{T}_i$ contains $\beta$ and $\alpha'$, where $\alpha'$ is the vertex of $H_i$ obtained by contracting $P_{i,j}$.  However, before $T_i$ is attached to $T_0$ each occurence of $\alpha'$ in $\mathcal{T}_i$ is replaced by $\alpha$, so some bag of $\mathcal{T}$ contains both $\alpha$ and $\beta$. Therefore $\mathcal{T}$ is a tree decomposition of $H$.

  Next we verify that, when $k=3$ exactly one bag of $\mathcal{T}$ contains $\{x_1,\ldots,x_3\}$.  The bag $B_{y_0}$ contains $x_1,\ldots,x_k$. When $k=3$, each neighbour of $z_i$ of $y_0$ has a bag that contains at most two of $\{x_1,x_2,x_3\}$.  Therefore, when $k=3$, $B_y$ is the unique bag in $\mathcal{T}$ that contains $\{x_1,x_2,x_3\}$.

	All that remains is to show that any triple of vertices of $H$ is contained in at most two bags of $\mathcal{T}$. By the inductive hypothesis, the only triples we need to be concerned about are the $3$-element subsets of  $\{x_0,x_1,\ldots,x_k\}$. The case $k=1$ is trivial since no triples can be formed by $\{x_0,x_1\}$.  When $k=2$, $B_{y_0}=\{x_0,x_1,x_2\}$.  In this case, the choice of $\tau$ ensures that there is at most one face $F_i$ of $M$ that has vertices of $S$, $P_1$, and $P_2$ on its boundary.  (Refer to \cref{boring_figure} to convince yourself of this.)  By the inductive hypothesis (applied to $H_i$) the bag $B_{z_i}$ is the only bag of $\mathcal{T}_i$ that contains $\{x_0,x_1,x_2\}$.  Therefore there are at most two bags, $B_y$ and $B_{z_i}$ of $\mathcal{T}$ that contain $\{x_0,x_1,x_2\}$.  This establishes (iii) for $k\in\{1,2\}$ so, from this point on we assume that $k=3$.

	We claim that, for each distinct $a,b\in \{1,2,3\}$ there is at most one face $F_i$ such that $V(F_i)\cap S_{x_j}\neq\emptyset$ for each $j\in\{0,a,b\}$.  Suppose to the contrary that there are two such faces $F_{i_1}$ and $F_{i_2}$.  Create a graph $M^+$ from $M$ by placing a vertex $v_{i_b}$ inside $F_{i_b}$ and adjacent to each vertex in $V(F_{i_b})$ for each $b\in\{1,2\}$.  Since $M$ is planar and $F_{i_1}$ and $F_{i_2}$ are distinct faces of $M$, $M^+$ is planar.  However, contracting each of $S_{x_0},\ldots,S_{x_3}$ in $M^+$ produces a graph isomorphic to the complete bipartite graph $K_{3,3}$, a contradiction.  (In this contracted graph, $x_0,x_a,x_b$ are on one side of the bipartition and $v_{i_a}$, $v_{i_b}$, and the node in $\{x_1,x_2,x_3\}\setminus\{x_a,x_b\}$ is on the other side.)

	We have already established that, when $k=3$, the only bag that contains $\{x_1,x_2,x_3\}$ is $B_{y_0}$, so we need only consider triples $\{x_0,x_a,x_b\}$ for distinct $a,b\in\{1,2,3\}$. By the preceding claim, there is at most one face $F_i$ with vertices of $S$, $P_a$, and $P_b$ on its boundary.  By the inductive hypothesis, the only other bag in $\mathcal{T}$ that contains $\{x_0,x_a,x_b\}$ is the bag $B_{z_i}$.  This completes the proof.
\end{proof}

@@

	Use the layering $\mathcal{L}=\langle L_0',L_1'\ldots\rangle$ where $L_i'=L_{\lfloor d/2\rfloor i}\cup \cdots \cup L_{\lfloor d/2\rfloor(i+1)-1}$ for each integer $i\ge 0$. That this is a layering of $G$ follows from the fact that $\dist_{G'}(v,w)\le \lfloor d/2\rfloor$ for each edge $vw\in E(G)$.  Since $|L_j\cap Y_x|\le 3$ for each integer $j\ge 0$, $|L_i'\cap Y_x|\le 3\lfloor d/2\rfloor$ and therefore $|L_i'\cap S_x|= |L_i'\cap Y_x|+ |L_i'\cap X_x|\le \lfloor d/2\rfloor + d -3$ for each integer $i\ge 0$ and each $x\in V(H)$. The result now follows from \cref{PartitionProduct}.
\end{proof}
